---
title: "Biostatistical Analysis of Ethanol Pharmacokinetics"
author: "Anna Vronskaia, Alexey Gordeev, Ekaterina Danilina, Milana Sagitova, Sergey Tsaregorodtsev"
date: "2026-01-31"
output:
  html_document:
    number_sections: false
    code_folding: show
    toc: true
    toc_float: true
    toc_depth: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(flextable))
library(patchwork)
library(GGally)
library(easystats)
library(regressinator)
library(broom)
library(gridExtra)
library(grid)
library(ggfortify)
library(performance)
library(ggeffects)
library(jtools)
library(sjPlot)
library(performance)
suppressPackageStartupMessages(library(lmtest))
suppressPackageStartupMessages(library(ggResidpanel))
library(sandwich)
library(ggeffects)
library(jtools)
library(gtsummary)
library(ggmosaic)
library(performance)
library(ggResidpanel)
library(table1)
library(ggfortify)
library(see)
```

```{r data_upload}
Jones_1984 <- read.csv("data/raw/Jones-1984.csv")
Ammon_1996 <- read.csv("data/raw/Ammon-1996.csv")
Yelland_2008 <- read.csv("data/raw/Yelland-2008.csv")
Emberger_2023 <- read.csv("data/raw/Thierauf-Emberger-2023.csv")
Wilkinson_IV <- read.csv("data/raw/Wilkinson-1976-IV.csv")
Wilkinson_PO <- read.csv("data/raw/Wilkinson-1976-PO.csv")
Vestal_1977_elderly <- read.csv("data/raw/Vestal-1977-elderly.csv")
Vestal_1977_young <- read.csv("data/raw/Vestal-1977-young.csv")

Kozawa_2007 <- read.csv("data/raw/Kozawa-2007.csv")
Eriksson_1977 <- read.csv("data/raw/Eriksson-1977.csv")
```

```{r theme_set}
standard_theme <- theme_bw() +
  theme(plot.title = element_text(size = 30, hjust = 0.5),
    plot.subtitle = element_text(size = 25, hjust = 0.5),
    plot.caption = element_text(size = 20),
    axis.title = element_text(size = 25),
    axis.text.y = element_text(size = 20),
    axis.text.x = element_text(size = 20),
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 22),
    strip.text = element_text(size = 22))
theme_set(standard_theme)
set_flextable_defaults(table.layout = "autofit")
flex_default_fun <- function(x) {
  x %>% 
  theme_box() %>%
  font(fontname = "Times New Roman", part = "all") %>% 
  fontsize(size = 13, part = "header") %>% 
  fontsize(size = 12, part = "body") %>% 
  flextable::align(
  i = 1,
  j = NULL,
  align = "center",
  part = "header")
}
```

# 1. Human data

## 1.1 Data unification

```{r ammon_clean}
Ammon_1996_clean <- Ammon_1996 %>% 
  mutate(case = paste(subject_id, occasion, sep = "-")) %>% 
  mutate(route = route %>% str_remove_all(" infusion")) %>% 
  rename(infusion_duration_min = duration_inf_min)
rm(Ammon_1996)
```

```{r jones_clean}
Jones_1984_clean <- Jones_1984 %>% 
  mutate(subject_id = paste("J", subject_id, sep = "-"),
         sex = case_when(sex == "male" ~ "M",
                         sex == "female" ~ "F"))
rm(Jones_1984)
```

```{r yelland_clean}
Yelland_2008_clean <- Yelland_2008 %>% 
  mutate(case = paste(subject_id, occasion, sep = "-"))
rm(Yelland_2008)
```

```{r emberger_clean}
Emberger_2023_clean <- Emberger_2023 %>% 
  filter(!is.na(conc_g_per_L)) %>% 
  rename(wt_kg = WT_kg,
         sex = Sex,
         age = Age,
         bh_cm = BH_cm,
         bmi = BMI) %>% 
  mutate(subject_id = str_c("TE", "-", subject_id))
rm(Emberger_2023)
```

```{r wilkinson_iv_clean}
Wilkinson_IV_clean <- Wilkinson_IV %>% 
  filter(!is.na(conc_g_per_L)) %>% 
  mutate(route = route %>% str_remove_all(" infusion")) %>% 
  rename(wt_kg = WT_kg,
         sex = Sex)
rm(Wilkinson_IV)
```

```{r wilkinson_po_clean}
Wilkinson_PO_clean <- Wilkinson_PO %>% 
  filter(!is.na(conc_g_per_L)) %>% 
  mutate(case = paste(subject_id, occasion, sep = "-")) %>% 
  rename(wt_kg = WT_kg,
         sex = Sex,
         bh_cm = BH_cm,
         bsa_m2 = BSA_m2,
         age = Age)  
rm(Wilkinson_PO)
```

```{r westal_old_clean}
Vestal_1977_e_pk <- Vestal_1977_elderly %>% 
rename_with(tolower) %>% 
  rename(Cmax = cmax_mg_dl) %>% 
  mutate(Cmax = Cmax / 100,
         route = case_when(
           str_detect(str_to_lower(route), "iv") ~ "IV",
           str_detect(str_to_lower(route), "po") ~ "PO"),
         subject_id = str_c("V-E-", subject_id))
rm(Vestal_1977_elderly)
```

```{r vestal_young_clean}
Vestal_1977_y_pk <- Vestal_1977_young %>% 
rename_with(tolower) %>% 
  rename(Cmax = cmax_mg_dl) %>% 
  mutate(Cmax = Cmax / 100,
         route = case_when(
           str_detect(str_to_lower(route), "iv") ~ "IV",
           str_detect(str_to_lower(route), "po") ~ "PO"),
         subject_id = str_c("V-Y-", subject_id))  
rm(Vestal_1977_young)
```

## 1.2. Data plots

### 1.2.1. Individual curves concentration-time

```{r ammon_ind_plot, fig.width=24, fig.height=16, dpi=300}
Ammon_1996_clean %>% 
  select(subject_id, conc_g_per_L, time_min, occasion) %>% 
  mutate(occasion = occasion %>% as.factor) %>% 
  ggplot(aes(color = occasion)) +
  geom_line(aes(x = time_min, y = conc_g_per_L, group = occasion), linewidth = 2) +
  facet_wrap(subject_id~.)+
  scale_x_continuous(breaks = seq(0, 420, by = 50))+
  scale_y_continuous(breaks = seq(0, 1, by = 0.25))+
    labs(title = "Concentration-time plot", subtitle = "Individual curves for each subject", x = "Time, min", y = "Concentration, g/L", color = "Визит", caption = "Source: Ammon-1996")+
  theme(legend.position = "top")
```

```{r jones_ind_plot, fig.height=12, fig.width=15, dpi=300}
ggplot(Jones_1984_clean, aes(x = time_min, y = conc_g_per_L)) +
  geom_line(aes(group = subject_id)) +
  scale_y_continuous(breaks = seq(0, 1.4, 0.1)) +
  scale_x_continuous(breaks = seq(0, 420, 30)) +
  # coord_cartesian(xlim = c(60, 420)) +
  labs(title = "Concentration-time plot",
       subtitle = "Individual curves for each subject",
       caption = "Source: Jones-1984",
       x = "Time, min",
       y = "Concentration, g/L")
```

```{r yelland_ind_plot, fig.width=24, fig.height=16, dpi=300}
Yelland_2008_clean %>% 
  select(subject_id, conc_g_per_L, time_min, occasion) %>% 
  mutate(occasion = occasion %>% as.factor) %>% 
  ggplot(aes(color = occasion)) +
  geom_line(aes(x = time_min, y = conc_g_per_L, group = occasion), linewidth = 2) +
  facet_wrap(subject_id~.)+
  scale_x_continuous(breaks = seq(0, 420, by = 50))+
  scale_y_continuous(breaks = seq(0, 1, by = 0.25))+
    labs(title = "Concentration-time plot", subtitle = "Individual curves for each subject", x = "Time, min", y = "Concentration, g/L", color = "Occasion", caption = "Source: Yelland-2008")+
  theme(legend.position = "top")
```

```{r emberger_ind_plot, fig.width=24, fig.height=16}
Emberger_2023_clean %>% 
  select(subject_id, conc_g_per_L, time_min) %>% 
  mutate(subject_id = subject_id %>% str_remove_all("TE-")) %>%
  ggplot(aes(color = subject_id)) +
  geom_line(aes(x = time_min, y = conc_g_per_L, group = subject_id), linewidth = 2) +
  scale_x_continuous(breaks = seq(0, 420, by = 50)) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.25)) +
  scale_color_discrete(breaks = as.character(1:10)) +
    labs(title = "Concentration-time plot",
         subtitle = "Individual curves for each subject", 
         x = "Time, min", y = "Concentration, g/L",
         color = "Subject", 
         caption = "Source: Thierauf-Emberger-2023") +
  theme(legend.position = "top")
```

```{r wilkinson_iv_ind_plot, fig.height=12, fig.width=15, dpi=300}
ggplot(Wilkinson_IV_clean, aes(x = time_min, y = conc_g_per_L)) +
  geom_line(aes(group = subject_id)) +
  scale_y_continuous(breaks = seq(0, 1.4, 0.1)) +
  scale_x_continuous(breaks = seq(0, 510, 30)) +
  # coord_cartesian(xlim = c(60, 420)) +
  labs(title = "Concentration-time plot",
       subtitle = "Individual curves for each subject",
       caption = "Source: Wilkinson-1976-IV",
       x = "Time, min",
       y = "Concentration, g/L")
```

```{r wilkinson_po_ind_plot, fig.width=20, fig.height=14}
Wilkinson_PO_clean %>%
  select(subject_id, conc_g_per_L, time_min, occasion) %>% 
  mutate(occasion = occasion %>% as.factor) %>% 
  ggplot(aes(color = occasion)) +
  geom_line(aes(x = time_min, y = conc_g_per_L, group = occasion), linewidth = 2) +
  facet_wrap(subject_id~.)+
  scale_x_continuous(breaks = seq(0, 420, by = 50))+
  scale_y_continuous(breaks = seq(0, 1, by = 0.25))+
    labs(title = "Concentration-time plot", subtitle = "Individual curves for each subject", x = "Time, min", y = "Concentration, g/L", color = "Occasion", caption = "Source: Wilkinson-1976-PO")+
  theme(legend.position = "top")
```

### 1.2.2. Mean и standard deviation of concentration

In each time point time is averaged.

```{r ammon_mean_sd_plot, fig.width=24, fig.height=16, dpi=300, message=FALSE}
Ammon_1996_clean %>% 
  group_by(case) %>%
  mutate(point_id = row_number() %>% as.factor) %>% 
  ungroup() %>%
  mutate(occasion = occasion %>% as.factor) %>% 
  group_by(occasion, point_id) %>% 
  summarise(conc_avg = mean(conc_g_per_L), conc_max = mean(conc_g_per_L) + sd(conc_g_per_L), conc_min = mean(conc_g_per_L) - sd(conc_g_per_L), time_min_avg = mean(time_min)) %>% 
  ungroup() %>% 
  ggplot() +
  geom_line(aes(x = time_min_avg, y = conc_avg, group = occasion, color = occasion)) +
  geom_errorbar(aes(x = time_min_avg, ymin = conc_min, ymax = conc_max), width = 6, linewidth = 0.5) +
  scale_x_continuous(breaks = seq(0, 400, by = 50)) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.25)) +    
  labs(title = "Concentration-time plot",
       subtitle = "Mean ± SD",
       x = "Time, min",
       y = "Concentration, g/L",
       color = "Occasion",
       caption = "Source: Ammon-1996")
```

```{r jones_mean_sd_plot, fig.height=12, fig.width=15, dpi=300}
Jones_1984_clean %>% 
  group_by(subject_id) %>%
  mutate(point_id = row_number() %>% as.factor) %>% 
  ungroup() %>% 
  group_by(point_id) %>% 
  summarise(conc_avg = mean(conc_g_per_L), conc_max = mean(conc_g_per_L) + sd(conc_g_per_L), conc_min = mean(conc_g_per_L) - sd(conc_g_per_L), time_min_avg = mean(time_min)) %>% 
  ggplot() +
  geom_line(aes(x = time_min_avg, y = conc_avg)) +
  geom_errorbar(aes(x = time_min_avg, ymin = conc_min, ymax = conc_max), width = 6, linewidth = 0.5) +
  scale_x_continuous(breaks = seq(0, 400, by = 50)) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.25)) +    
  labs(title = "Concentration-time plot",
       subtitle = "Mean ± SD",
       x = "Time, min",
       y = "Concentration, g/L",
       caption = "Source: Jones-1984")
```

```{r yelland_mean_sd_plot, fig.width=16, fig.height=9, dpi=300, message=FALSE}
Yelland_2008_clean %>%
  group_by(case) %>%
  mutate(point_id = row_number() %>% as.factor) %>% 
  ungroup() %>%
  mutate(occasion = occasion %>% as.factor) %>% 
  group_by(occasion, point_id) %>% 
  summarise(conc_avg = mean(conc_g_per_L), conc_max = mean(conc_g_per_L) + sd(conc_g_per_L), conc_min = mean(conc_g_per_L) - sd(conc_g_per_L), time_min_avg = mean(time_min)) %>% 
  ungroup() %>% 
  ggplot() +
  geom_line(aes(x = time_min_avg, y = conc_avg, group = occasion, color = occasion)) +
  geom_errorbar(aes(x = time_min_avg, ymin = conc_min, ymax = conc_max), width = 6, linewidth = 0.5) +
  scale_x_continuous(breaks = seq(0, 400, by = 50)) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.25)) +    
  labs(title = "Concentration-time plot",
       subtitle = "Mean ± SD",
       x = "Time, min",
       y = "Concentration, g/L",
       color = "Occasion",
       caption = "Source: Yelland-2008")
```

```{r emberger_mean_sd_plot, fig.width=24, fig.height=16, dpi=300}
Emberger_2023_clean %>% 
  group_by(subject_id) %>%
  mutate(point_id = row_number() %>% as.factor) %>% 
  ungroup() %>% 
  group_by(point_id) %>% 
  summarise(conc_avg = mean(conc_g_per_L), conc_max = mean(conc_g_per_L) + sd(conc_g_per_L), conc_min = mean(conc_g_per_L) - sd(conc_g_per_L), time_min_avg = mean(time_min)) %>% 
  ggplot() +
  geom_line(aes(x = time_min_avg, y = conc_avg)) +
  geom_errorbar(aes(x = time_min_avg, ymin = conc_min, ymax = conc_max), width = 6, linewidth = 0.5) +
  scale_x_continuous(breaks = seq(0, 400, by = 50)) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.25)) +    
  labs(title = "Concentration-time plot",
       subtitle = "Mean ± SD",
       x = "Time, min",
       y = "Concentration, g/L",
       caption = "Source: Thierauf-Emberger-2023")
```

```{r wilkinson_iv_mean_sd_plot, fig.width=24, fig.height=16, dpi=300}
Wilkinson_IV_clean %>% 
  group_by(subject_id) %>%
  mutate(point_id = row_number() %>% as.factor) %>% 
  ungroup() %>% 
  group_by(point_id) %>% 
  summarise(conc_avg = mean(conc_g_per_L), conc_max = mean(conc_g_per_L) + sd(conc_g_per_L), conc_min = mean(conc_g_per_L) - sd(conc_g_per_L), time_min_avg = mean(time_min)) %>% 
  filter(as.numeric(point_id) < 27) %>% 
  ggplot() +
  geom_line(aes(x = time_min_avg, y = conc_avg)) +
  geom_errorbar(aes(x = time_min_avg, ymin = conc_min, ymax = conc_max), width = 6, linewidth = 0.5) +
  scale_x_continuous(breaks = seq(0, 400, by = 50)) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.25)) +    
  labs(title = "Concentration-time plot",
       subtitle = "Mean ± SD",
       x = "Time, min",
       y = "Concentration, g/L",
       caption = "Source: Wilkinson-1976-IV")  
```

```{r wilkinson_po__mean_sd_plot, fig.width=24, fig.height=16, dpi=300}
Wilkinson_PO_clean %>%
  group_by(case) %>%
  mutate(point_id = row_number() %>% as.factor) %>% 
  ungroup() %>% 
  group_by(point_id) %>% 
  summarise(conc_avg = mean(conc_g_per_L), conc_max = mean(conc_g_per_L) + sd(conc_g_per_L), conc_min = mean(conc_g_per_L) - sd(conc_g_per_L), time_min_avg = mean(time_min)) %>% 
  ggplot() +
  geom_line(aes(x = time_min_avg, y = conc_avg)) +
  geom_errorbar(aes(x = time_min_avg, ymin = conc_min, ymax = conc_max), width = 6, linewidth = 0.5) +
  scale_x_continuous(breaks = seq(0, 400, by = 50)) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.25)) +    
  labs(title = "Concentration-time plot",
       subtitle = "Mean ± SD",
       x = "Time, min",
       y = "Concentration, g/L",
       caption = "Source: Wilkinson-1976-PO")  
```

## 1.3. Evaluation of PK-parameters

```{r pk_calc_functions}
#k is how much we add to the tmax index
pharma_fun <- function(conc, t, dose, k = 1) {
  # Modeling for calculation beta, C0, Vd 0 order
  t_sample <- {{t}}[(first(which.max({{conc}})) + k):length({{t}})]
  conc_sample <- conc[(first(which.max({{conc}})) + k):length({{conc}})]
  dose <- first({{dose}})
  model <- lm(conc_sample ~ t_sample)
  Beta <- -as.numeric(model$coefficients[2])
  C_0_0_order <- as.numeric(model$coefficients[1])
  V_d_0_order <- dose / C_0_0_order
  
  # Modeling for calculation kel, C0, Vd 1 order
  # indexes <- conc_sample > 0
  # model_1 <- lm(log(conc_sample[indexes]) ~ t_sample[indexes])
  # Kel <- -as.numeric(model_1$coefficients[2])
  # C_0_1_order <- exp(as.numeric(model_1$coefficients[1]))
  # V_d_1_order <- dose / C_0_1_order
  
  # Calculation of AUC0-t
  
  # if (length({{t}}) < 2) AUC <- NA
  # if (anyNA({{t}}) | anyNA({{conc}})) AUC <- NaN
  t0 = {{t}}[-length({{t}})]
  t1 = {{t}}[-1]
  c0 = {{conc}}[-length({{conc}})]
  c1 = {{conc}}[-1]
  AUC = sum(((c0 + c1)/2)*(t1 - t0))
  
  # Calculation of Cmax
  cmax <- max(conc)
  
  # Calculation of Tmax
  tmax <- t[first(which.max(conc))]
  
  # Calculation of CL (roughly)
  cl = dose/AUC #l/(kg*min)
  # V = CL/Kel 
  
  return(list(Beta = Beta,
              # Kel = Kel,
              C_0_0_order = C_0_0_order,
              V_d_0_order = V_d_0_order,
              # C_0_1_order = C_0_1_order,
              # V_d_1_order = V_d_1_order,
              AUC = AUC,
              Cmax = cmax,
              Tmax = tmax,
              CL = cl))
}

stat_ph <- list(
        "Mean" = ~mean(., na.rm = TRUE) %>% round(4),
        "Median" = ~median(., na.rm = TRUE)%>% round(4),
        "Standard deviation"   = ~sd(., na.rm = TRUE) %>% round(4),
        "Coefficient of variation (CV), %" = ~round((sd(., na.rm = TRUE))/(mean(., na.rm = TRUE)) * 100, 2)
        )

```

```{r jones_pk}
Jones_1984_pk <- Jones_1984_clean %>% 
  arrange(time_min) %>% 
  nest(.by = subject_id, .key = "subject_data") %>% 
  mutate(pharma = purrr::map(subject_data,
                            ~pharma_fun(.$conc_g_per_L,
                                        .$time_min,
                                        .$dose_g_per_kg))) %>% 
  unnest_wider(col = pharma) %>% 
  select(-subject_data)

Jones_1984_pk <- Jones_1984_clean %>% 
  select(-c(time_min, conc_g_per_L)) %>% 
  group_by(subject_id) %>% 
  slice(1) %>% 
  ungroup() %>% 
  right_join(Jones_1984_pk)
rm(Jones_1984_clean)
```

```{r ammon_pk}
Ammon_1996_pk <- Ammon_1996_clean %>% 
  arrange(time_min) %>% 
  nest(.by = case, .key = "subject_data") %>% 
  mutate(pharma = purrr::map(subject_data,
                            ~pharma_fun(.$conc_g_per_L,
                                        .$time_min,
                                        .$dose_g_per_kg))) %>% 
  unnest_wider(col = pharma) %>% 
  select(-subject_data)

Ammon_1996_pk <- Ammon_1996_clean %>%
  select(-c(time_min, conc_g_per_L)) %>%
  group_by(case) %>%
  slice(1) %>%
  ungroup() %>%
  right_join(Ammon_1996_pk, by = join_by(case)) %>% 
  mutate(subject_id = str_c("A-IV-", subject_id)) %>% 
  select(-case)
rm(Ammon_1996_clean)
```

```{r yelland_pk}
Yelland_2008_pk <- Yelland_2008_clean %>% 
  arrange(time_min) %>% 
  nest(.by = case, .key = "subject_data") %>% 
  mutate(pharma = purrr::map(subject_data,
                            ~pharma_fun(.$conc_g_per_L,
                                        .$time_min,
                                        .$dose_g_per_kg))) %>% 
  unnest_wider(col = pharma) %>%
  select(-subject_data)

Yelland_2008_pk <- Yelland_2008_clean %>%
  select(-c(time_min, conc_g_per_L)) %>%
  group_by(case) %>%
  slice(1) %>%
  ungroup() %>%
  right_join(Yelland_2008_pk, by = join_by(case)) %>%
  mutate(subject_id = str_c("Y-", subject_id)) %>% 
  select(-case)
rm(Yelland_2008_clean)
```

```{r emberger_pk}
Emberger_2023_pk <- Emberger_2023_clean %>% 
  arrange(time_min) %>% 
  nest(.by = subject_id, .key = "subject_data") %>% 
  mutate(pharma = purrr::map(subject_data,
                            ~pharma_fun(.$conc_g_per_L,
                                        .$time_min,
                                        .$dose_g_per_kg))) %>% 
  unnest_wider(col = pharma) %>% 
  select(-subject_data)

Emberger_2023_pk <- Emberger_2023_clean %>% 
  select(-c(time_min, conc_g_per_L)) %>% 
  group_by(subject_id) %>% 
  slice(1) %>% 
  ungroup() %>% 
  right_join(Emberger_2023_pk)
rm(Emberger_2023_clean)
```

```{r wilkinson_iv_pk}
Wilkinson_IV_pk <- Wilkinson_IV_clean %>% 
  arrange(time_min) %>% 
  nest(.by = subject_id, .key = "subject_data") %>% 
  mutate(pharma = purrr::map(subject_data,
                            ~pharma_fun(.$conc_g_per_L,
                                        .$time_min,
                                        .$dose_g_per_kg))) %>% 
  unnest_wider(col = pharma) %>% 
  select(-subject_data)

Wilkinson_IV_pk <- Wilkinson_IV_clean %>% 
  select(-c(time_min, conc_g_per_L)) %>% 
  group_by(subject_id) %>% 
  slice(1) %>% 
  ungroup() %>% 
  right_join(Wilkinson_IV_pk) %>% 
  mutate(subject_id = str_c("W-IV-", subject_id))
rm(Wilkinson_IV_clean)
```

```{r wilkinson_po_pk}
Wilkinson_PO_pk <- Wilkinson_PO_clean %>% 
  arrange(time_min) %>% 
  nest(.by = case, .key = "subject_data") %>% 
  mutate(pharma = purrr::map(subject_data,
                            ~pharma_fun(.$conc_g_per_L,
                                        .$time_min,
                                        .$dose_g_per_kg))) %>% 
  unnest_wider(col = pharma) %>% 
  select(-subject_data)

Wilkinson_PO_pk <- Wilkinson_PO_clean %>%
  select(-c(time_min, conc_g_per_L)) %>%
  group_by(case) %>%
  slice(1) %>%
  ungroup() %>%
  right_join(Wilkinson_PO_pk, by = join_by(case)) %>% 
  mutate(subject_id = str_c("W-PO-", subject_id)) %>% 
  select(-case)
rm(Wilkinson_PO_clean)
```

## 1.4. Binding datasets

```{r join-all}
main_df <- bind_rows(Ammon_1996_pk,
                     Emberger_2023_pk,
                     Jones_1984_pk,
                     Wilkinson_IV_pk,
                     Wilkinson_PO_pk,
                     Yelland_2008_pk,
                     Vestal_1977_e_pk,
                     Vestal_1977_y_pk)

main_df <- main_df %>% 
  mutate(across(c(where(is.character), occasion, infusion_duration_min, -subject_id), ~as.factor(.x)))
```

```{r zero-first-orders-df}
# Only mean values by visits (the assumption of independence of observations is fulfilled)
WILK <- main_df %>% 
  filter(study_id == "Wilkinson-1976-PO") %>% 
  arrange(desc(dose_g_per_kg), .by_group = TRUE) %>% 
  group_by(subject_id) %>% 
  slice(1)
order_0_df <- main_df %>%
  filter(study_id != "Wilkinson-1976-PO") %>%
  group_by(subject_id) %>% 
  mutate(across(c(Beta, C_0_0_order, V_d_0_order, AUC, Cmax, Tmax), ~mean(.x))) %>%
  slice(1) %>% 
  ungroup() %>% 
  select(-occasion) %>% 
  bind_rows(WILK) %>% 
  select(-c(
    # Kel, 
    occasion #,
    # contains("1_order")
    )) 

# nrow(order_0_df)
# length(unique(order_0_df$subject_id))
# order_1_df <- main_df %>% filter(Cmax < 0.23)
```

## 1.5. Exploratory analysis

### 1.5.1. Tables

```{r}
main_df %>% 
  select(-subject_id) %>% 
  gtsummary::tbl_summary()
```

```         
1.  PK-parameters (endpoints)
•   AUC, Cmax, 
•   C0_0, C0_1
•   Beta, Kel, 
•   Tmax
•   Vd_0,Vd_1
•   CL
2. Experimental conditions
•   Dose (g/kg)
•   Route (IV/PO)
•   Occasion (1/2/NA)
•   Food (fed/fasted)
•   Infusion_duration_min
•   Study_id
3.  Patient's personal data
•   WT, BH, BMI, BSA, LBM
•   Sex, Age
```

```{r}
label(main_df$study_id)   <- "Study"
label(main_df$sex)   <- "Sex"
label(main_df$dose_g_per_kg)    <- "Dose"
label(main_df$food) <- "Food"
label(main_df$site) <- "Blood sampling"
label(main_df$age) <- "Age"
label(main_df$route) <- "Route"

units(main_df$dose_g_per_kg)   <- "g per kg"
units(main_df$age)   <- "years"

tbl_output <- table1(~ study_id + dose_g_per_kg + food + site + age + sex|route, data = main_df, render.missing = NULL, overall=c(left="Total"))


tbl_output
```

```{r distribution-plots, error=FALSE, warning=FALSE, fig.width=12, fig.height=8, dpi=300}

main_df_long <- main_df %>% 
  select(where(is.numeric)) %>% 
  pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "value"
  )

ggplot(main_df_long, aes(x = value)) +
  geom_histogram(bins = 30) +
  facet_wrap(~ variable, scales = "free") +
  theme_bw() +
  labs(x = NULL, y = NULL)
```

```{r density-plots, fig.width=12, fig.height=8, dpi=300}
ggplot(main_df_long, aes(x = value)) +
  geom_density(na.rm = TRUE) +
  facet_wrap(~ variable, scales = "free") +
  theme_bw() +
  labs(x = NULL, y = NULL)
```

We can take the logarithm of variables with right-sided skewness (V_d_0, AUC, CL)

```{r density_plot_log, fig.width=12, fig.height=8, dpi=300}

main_df_log_long <- main_df %>% 
  mutate(across(c(
    # Kel,
    V_d_0_order,
    # V_d_1_order,
    # C_0_1_order,
    AUC,
    CL), function(x) log(x))) %>% 
  select(where(is.numeric)) %>% 
  pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "value"
  )

ggplot(main_df_log_long, aes(x = value)) +
  geom_density(na.rm = TRUE) +
  facet_wrap(~ variable, scales = "free") +
  theme_bw() +
  labs(x = NULL, y = NULL)
```

```{r factor-variables-distribution}
main_df_fac_long <- main_df %>% 
  select(where(is.factor)) %>% 
  pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "value"
  )

ggplot(main_df_fac_long, aes(x = value)) +
  geom_bar() +
  facet_wrap(~ variable, scales = "free_x") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = NULL, y = NULL)

```

### 1.5.2. Boxplots

```{r theme_custom}
theme_custom <- theme(
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    axis.title = element_text(size = 14),
    plot.title = element_text(size = 20, hjust = 0.5),
    plot.subtitle = element_text(size = 16, hjust = 0.5),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14),
    panel.background = element_rect(fill = "white"), 
    panel.grid = element_blank()
  )

gender_colours <- c(
      "M" = "#8694AC",
      "F" = "#f28482"
    )

route_colours <- c(
      "IV" = "#DDB892",
      "PO" = "#BAC893"
    )

food_colours <- c(
      "fed" = "#FFB35C",
      "fasted" = "#CDE8F4"
    )

site_colours <- c(
      "capillary" = "#FE3448",
      "venous" = "#7C7F65"
    )

gender_fill <- scale_fill_manual(
    name = "Sex",
    values = gender_colours
  )

route_fill <- scale_fill_manual(
    name = "Route",
    values = route_colours
  )

food_fill <- scale_fill_manual(
    name = "Food",
    values = food_colours
  )

site_fill <- scale_fill_manual(
    name = "Site",
    values = site_colours
  )


gender_color <- scale_colour_manual(
    name = "Sex",
    values = gender_colours
  )
```

```{r, warning=FALSE, message=FALSE, fig.width=12, fig.height=5}
p1 <- main_df %>% 
  filter(!is.na(sex)) %>% 
  ggplot(aes(x = sex, y = Tmax)) +
  geom_violin(aes(fill = sex)) +
  geom_boxplot(width = 0.1, color = "black", alpha = 0.5)+
  geom_jitter(color = "gray40", alpha = 0.5, size = 1.5)+
  theme(
    legend.position = "none"
  ) + 
  theme_custom + 
  gender_fill

p2 <- main_df %>% 
  filter(!is.na(sex)) %>%
  ggplot()+
  geom_histogram(aes(x = Tmax, fill = sex), colour = "gray40") + 
  theme_custom + 
  gender_fill

p1 + p2 + plot_annotation(title = "Distribution of Tmax by gender", theme = theme(plot.title = element_text(size = 20)))
  
```

```{r, warning=FALSE, message=FALSE, fig.width=12, fig.height=5}
p3 <- main_df %>% 
  filter(!is.na(sex)) %>%
  ggplot(aes(x = sex, y = Cmax)) +
  geom_violin(aes(fill = sex)) +
  geom_boxplot(width = 0.1, color = "black", alpha = 0.5)+
  geom_jitter(color = "gray40", alpha = 0.5, size = 1.5)+
  theme(
    legend.position = "none"
  ) + 
  theme_custom + 
  gender_fill

p4 <-  main_df %>% 
  filter(!is.na(sex)) %>%
  ggplot()+
  geom_histogram(aes(x = Cmax, fill = sex), colour = "gray40") + 
  theme_custom + 
  gender_fill

p3 + p4 + plot_annotation(title = "Distribution of Cmax by gender", theme = theme(plot.title = element_text(size = 20)))
  
```

```{r, warning=FALSE, message=FALSE, fig.width=12, fig.height=5}
p5 <- main_df %>% 
  filter(!is.na(sex)) %>% 
  ggplot(aes(x = sex, y = AUC)) +
  geom_violin(aes(fill = sex)) +
  geom_boxplot(width = 0.1, color = "black", alpha = 0.5)+
  geom_jitter(color = "gray40", alpha = 0.5, size = 1.5)+
  theme(
    legend.position = "none"
  ) + 
  theme_custom + 
  gender_fill

p6 <- main_df %>% 
  filter(!is.na(sex)) %>%
  ggplot()+
  geom_histogram(aes(x = AUC, fill = sex), colour = "gray40") + 
  theme_custom + 
  gender_fill

p5 + p6 + plot_annotation(title = "Distribution of AUC by gender", theme = theme(plot.title = element_text(size = 20)))
```

```{r, warning=FALSE, message=FALSE, fig.width=12, fig.height=5}
p7 <- main_df %>% 
  ggplot(aes(x = route, y = Tmax)) +
  geom_violin(aes(fill = route)) +
  geom_boxplot(width = 0.1, color = "black", alpha = 0.5)+
  geom_jitter(color = "gray40", alpha = 0.5, size = 1.5)+
  theme(
    legend.position = "none"
  ) + 
  theme_custom +
  route_fill

p8 <- main_df %>% 
  ggplot()+
  geom_histogram(aes(x = Tmax, fill = route), colour = "gray40") + 
  theme_custom +
  route_fill

p7 + p8 + plot_annotation(title = "Distribution of Tmax by route", theme = theme(plot.title = element_text(size = 20)))
  
```

```{r, warning=FALSE, message=FALSE, fig.width=12, fig.height=5}
p9 <- main_df %>% 
  ggplot(aes(x = route, y = Cmax)) +
  geom_violin(aes(fill = route)) +
  geom_boxplot(width = 0.1, color = "black", alpha = 0.5)+
  geom_jitter(color = "gray40", alpha = 0.5, size = 1.5)+
  theme(
    legend.position = "none"
  ) + 
  theme_custom + 
  route_fill

p10 <-  main_df %>% 
  ggplot()+
  geom_histogram(aes(x = Cmax, fill = route), colour = "gray40") + 
  theme_custom + 
  route_fill

p9 + p10 + plot_annotation(title = "Distribution of Cmax by route", theme = theme(plot.title = element_text(size = 20)))
  
```

```{r, warning=FALSE, message=FALSE, fig.width=12, fig.height=5}
p11 <- main_df %>% 
  ggplot(aes(x = route, y = AUC)) +
  geom_violin(aes(fill = route)) +
  geom_boxplot(width = 0.1, color = "black", alpha = 0.5)+
  geom_jitter(color = "gray40", alpha = 0.5, size = 1.5)+
  theme(
    legend.position = "none"
  ) + 
  theme_custom + 
  route_fill

p12 <- main_df %>% 
  filter(!is.na(sex)) %>%
  ggplot()+
  geom_histogram(aes(x = AUC, fill = route), colour = "gray40") + 
  theme_custom + 
  route_fill

p11 + p12 + plot_annotation(title = "Distribution of AUC by route", theme = theme(plot.title = element_text(size = 20)))
```

```{r, warning=FALSE, message=FALSE, fig.width=12, fig.height=5}
p13 <- main_df %>% 
  ggplot(aes(x = food, y = Tmax)) +
  geom_violin(aes(fill = food)) +
  geom_boxplot(width = 0.1, color = "black", alpha = 0.5)+
  geom_jitter(color = "gray40", alpha = 0.5, size = 1.5)+
  theme(
    legend.position = "none"
  ) + 
  theme_custom +
  food_fill

p14 <- main_df %>% 
  ggplot()+
  geom_histogram(aes(x = Tmax, fill = food), colour = "gray40") + 
  theme_custom +
  food_fill

p13 + p14 + plot_annotation(title = "Distribution of Tmax by food consumption", theme = theme(plot.title = element_text(size = 20)))
  
```

```{r, warning=FALSE, message=FALSE, fig.width=12, fig.height=5}
p15 <- main_df %>% 
  ggplot(aes(x = food, y = Cmax)) +
  geom_violin(aes(fill = food)) +
  geom_boxplot(width = 0.1, color = "black", alpha = 0.5)+
  geom_jitter(color = "gray40", alpha = 0.5, size = 1.5)+
  theme(
    legend.position = "none"
  ) + 
  theme_custom + 
  food_fill

p16 <-  main_df %>% 
  ggplot()+
  geom_histogram(aes(x = Cmax, fill = food), colour = "gray40") + 
  theme_custom + 
  food_fill

p15 + p16 + plot_annotation(title = "Distribution of Cmax by food consumption", theme = theme(plot.title = element_text(size = 20)))
  
```

```{r, warning=FALSE, message=FALSE, fig.width=12, fig.height=5}
p17 <- main_df %>% 
  ggplot(aes(x = food, y = AUC)) +
  geom_violin(aes(fill = food)) +
  geom_boxplot(width = 0.1, color = "black", alpha = 0.5)+
  geom_jitter(color = "gray40", alpha = 0.5, size = 1.5)+
  theme(
    legend.position = "none"
  ) + 
  theme_custom + 
  food_fill

p18 <- main_df %>% 
  filter(!is.na(sex)) %>%
  ggplot()+
  geom_histogram(aes(x = AUC, fill = food), colour = "gray40") + 
  theme_custom + 
  food_fill

p17 + p18 + plot_annotation(title = "Distribution of AUC by food consumption", theme = theme(plot.title = element_text(size = 20)))
```

```{r, warning=FALSE, message=FALSE, fig.width=12, fig.height=5}
p19 <- main_df %>% 
  ggplot(aes(x = site, y = Tmax)) +
  geom_violin(aes(fill = site)) +
  geom_boxplot(width = 0.1, color = "black", alpha = 0.5)+
  geom_jitter(color = "gray40", alpha = 0.5, size = 1.5)+
  theme(
    legend.position = "none"
  ) + 
  theme_custom +
  site_fill

p20 <- main_df %>% 
  ggplot()+
  geom_histogram(aes(x = Tmax, fill = site), colour = "gray40") + 
  theme_custom +
  site_fill

p19 + p20 + plot_annotation(title = "Distribution of Tmax by blood sampling", theme = theme(plot.title = element_text(size = 20)))
  
```

```{r, warning=FALSE, message=FALSE, fig.width=12, fig.height=5}
p21 <- main_df %>% 
  ggplot(aes(x = site, y = Cmax)) +
  geom_violin(aes(fill = site)) +
  geom_boxplot(width = 0.1, color = "black", alpha = 0.5)+
  geom_jitter(color = "gray40", alpha = 0.5, size = 1.5)+
  theme(
    legend.position = "none"
  ) + 
  theme_custom + 
  site_fill

p22 <-  main_df %>% 
  ggplot()+
  geom_histogram(aes(x = Cmax, fill = site), colour = "gray40") + 
  theme_custom + 
  site_fill

p21 + p22 + plot_annotation(title = "Distribution of Cmax by blood sampling", theme = theme(plot.title = element_text(size = 20)))
  
```

```{r, warning=FALSE, message=FALSE, fig.width=12, fig.height=5}
p23 <- main_df %>% 
  ggplot(aes(x = site, y = AUC)) +
  geom_violin(aes(fill = site)) +
  geom_boxplot(width = 0.1, color = "black", alpha = 0.5)+
  geom_jitter(color = "gray40", alpha = 0.5, size = 1.5)+
  theme(
    legend.position = "none"
  ) + 
  theme_custom + 
  site_fill

p24 <-  main_df %>% 
  ggplot()+
  geom_histogram(aes(x = AUC, fill = site), colour = "gray40") + 
  theme_custom + 
  site_fill

p23 + p24 + plot_annotation(title = "Distribution of AUC by blood sampling", theme = theme(plot.title = element_text(size = 20)))
  
```

```{r Correlation, fig.height=8, fig.width=14, dpi = 200}
pk_num <- main_df %>% 
  select(where(is.numeric)) %>% 
  select(-lbm_kg, -bsa_m2, -bh_cm, -bmi) #лишние убрал


cor_mat <- cor(pk_num, use = "pairwise.complete.obs")

cor_mat %>% 
  ggcorr(., label = TRUE)
  
```

Testing hypotheses about differences between groups by category "food" and "site"

H0: mean (fasted) = mean(fed) H1: mean (fasted) != mean(fed)

```{r t-test в группах по food}
t.test(log(AUC) ~ food, data = order_0_df) 

t.test(log(Cmax) ~ food, data = order_0_df)

t.test(Tmax ~ food, data = order_0_df)

t.test(Beta ~ food, data = order_0_df)

t.test(C_0_0_order ~ food, data = order_0_df)

t.test(V_d_0_order ~ food, data = order_0_df)

```

H0: mean (capillary) = mean(venous) H1: mean (capillary) != mean(venous)

```{r t-test в группах по site}
t.test(log(AUC) ~ site, data = order_0_df) 

t.test(log(Cmax) ~ site, data = order_0_df)

t.test(Tmax ~ site, data = order_0_df)

t.test(Beta ~ site, data = order_0_df)

t.test(C_0_0_order ~ site, data = order_0_df)

t.test(V_d_0_order ~ site, data = order_0_df)
```

Testing hypotheses about differences between groups by category "route" and "sex".

```{r t-test 0_order route_sex}
t.test(Beta ~ route, data = order_0_df) 

t.test(Beta ~ sex, data = order_0_df)

t.test(V_d_0_order ~ route, data = order_0_df) 

t.test(V_d_0_order ~ sex, data = order_0_df)

t.test(log(AUC) ~ route, data = order_0_df) 

t.test(log(AUC) ~ sex, data = order_0_df)

t.test(Cmax ~ route, data = order_0_df)

t.test(Cmax ~ sex, data = order_0_df)

t.test(Tmax ~ route, data = order_0_df)

t.test(Tmax ~ sex, data = order_0_df)
```

**Results**: The administration route groups differed in all pharmacokinetic parameters. The gender groups differed in all pharmacokinetic parameters except Tmax.


## 1.6. Main Regression Analysis

### 1.6.1. Regression models for AUC

#### 1.6.1.1. Additive model with main variables (dose, route of administration, blood sampling method)

log(AUC) \~ dose_g_per_kg + route + site

```{r fit_lm}
fit_AUC_mod <- lm(log(AUC) ~ dose_g_per_kg + route + site, order_0_df) #+food
fit_AUC_submod <- lm(log(AUC) ~ 1, order_0_df)

#broom::tidy(fit_AUC_mod, conf.int = TRUE)
#broom::tidy(fit_AUC_submod, conf.int = TRUE)

summ(fit_AUC_mod, confint = TRUE, digits = 3)
#summary(fit_AUC_mod)
#summary(fit_AUC_submod)

anova(fit_AUC_mod, fit_AUC_submod)
```

```{r, fig.width=8, fig.height=10}
check_model(fit_AUC_mod)
```

```{r}
ggpredict(fit_AUC_mod, terms = c("dose_g_per_kg", "route", "site"), ci_level = 0) %>% plot(show_data = TRUE, log_y = FALSE, jitter = .01) + labs(subtitle = "Adjusted R-squared:  0.7626")
```

```{r, fig.width=7, fig.height=5, dpi=300}
plot_coefs(fit_AUC_mod, omit.coefs = NULL, scale = TRUE, ci_level = 0.95, plot.distributions = FALSE, inner_ci_level = .1, colors = "CUD") + labs(title = "Regression coeffecients with 95% CI")
```

```{r, include=FALSE}
table(order_0_df$site, order_0_df$food)
```

#### 1.6.1.2. Additive model with additional variables

log(AUC) \~ dose_g_per_kg + route + site + sex

```{r fit_lm_add}
fit_AUC_mod_add <- lm(log(AUC) ~ dose_g_per_kg + route + site + sex, order_0_df) #age cannot be added

summ(fit_AUC_mod, confint = TRUE, digits = 3)
summ(fit_AUC_mod_add, confint = TRUE, digits = 3)


# summary(fit_AUC_mod)
# summary(fit_AUC_mod_add)


#anova(fit_AUC_mod, fit_AUC_mod_add) Anova does not work with variable sex
```

```{r, fig.width=8, fig.height=10}
check_model(fit_AUC_mod_add)
```

```{r, include=FALSE}
order_0_df %>% filter(!is.na(age)) %>% select(dose_g_per_kg, site, route, age)
# dose 0.57 clearly corresponds to route IV
```

```{r}
ggpredict(fit_AUC_mod_add, terms = c("dose_g_per_kg", "route", "site", "sex")) %>% plot(show_data = TRUE, log_y = F, jitter = .01) #+ labs(subtitle = "Adjusted R-squared:  0.7626")
```

#### 1.6.1.3. Model with interactions

log(AUC) \~ dose_g_per_kg\*route + site

It is logical that `AUC` will depend on the combination of dose and method of administration.

```{r fit_lm_int}
fit_AUC_mod_int <- lm(log(AUC) ~ dose_g_per_kg*route + site, order_0_df)
# fit_AUC_mod_int2 <- lm(log(AUC) ~ dose_g_per_kg + route*site, order_0_df)
# fit_AUC_mod_int3 <- lm(log(AUC) ~ dose_g_per_kg*site + route, order_0_df)
# fit_AUC_mod_int4 <- lm(log(AUC) ~ dose_g_per_kg*site*route, order_0_df)


summ(fit_AUC_mod_int, confint = TRUE, digits = 3)
```

```{r}
anova(fit_AUC_mod_int, fit_AUC_submod)
```


```{r, fig.width=8, fig.height=10}
check_model(fit_AUC_mod_int)
```

```{r}
ggpredict(fit_AUC_mod_int, terms = c("dose_g_per_kg", "route", "site"), ci_level = 0) %>% plot(show_data = TRUE, log_y = FALSE, jitter = .01) + labs(subtitle = "Adjusted R-squared:  0.799") + xlab("Dose (g per kg)")
```

```{r, fig.width=7, fig.height=5, dpi=300}
plot_coefs(fit_AUC_mod_int, omit.coefs = NULL, scale = TRUE, ci_level = 0.95, plot.distributions = FALSE, inner_ci_level = .1, colors = "CUD") + labs(title = "Regression coeffecients with 95% CI")
```

```{r}
plot_model(fit_AUC_mod_int, show.values = TRUE,  show.intercept = TRUE, value.offset = 0.3,  axis.labels = c("Dose (g per kg)", "Route (PO)", "Blood sampling (venous)", "Dose (g per kg) x Route (PO)", "Intercept"), title = "Regression coeffecients with 95% CI") +
  theme_nice()
```

**Coclusions:**

Including the interaction between `dose_g_per_kg` and `route` increases $R^2$ by about 3%.

### 1.6.2. Regression models for Tmax

```{r, tmax-full}
order_0_df <- order_0_df %>% 
  mutate(bmi = ifelse(is.na(bmi) & !is.na(wt_kg) & !is.na(bh_cm),
                      wt_kg/(bh_cm/100)^2, bmi),
         bsa_m2 = ifelse(is.na(bsa_m2) & !is.na(wt_kg) & !is.na(bh_cm),
                         sqrt((bh_cm * wt_kg)/3600), bsa_m2))

t_max_full_m <- lm(Tmax ~ dose_g_per_kg * food + route + wt_kg, data = order_0_df)

summary(t_max_full_m)
```

```{r, fig.width=20, fig.height=15}
check_model(t_max_full_m)
```

```{r, fig.width=20, fig.height=15}
ggResidpanel::resid_xpanel(t_max_full_m)
```

```{r, fig.width=20, fig.height=15}
ggResidpanel::resid_panel(t_max_full_m, plots = c("lev", "cookd"))
```

Estimating model coefficients using sandwich estimator

```{r}
t_max_coef <- coeftest(t_max_full_m, vcov. = vcovHC, type = "HC3") %>%
broom::tidy(conf.int = TRUE) %>% 
  mutate(term = c("Intercept",
                  "Dose (g/kg)",
                  "Food status: \"Fed\"",
                  "Peroral route",
                  "Weight, kg",
                  "Interaction between\ndose and status \"fed\"")) %>% 
  rename("Variable" = term,
         "β estimate" = estimate,
         "Standard error" = std.error,
         "Wald t-test statistics" = statistic,
         "P-value" = p.value,
         "95%CI low" = conf.low,
         "95%CI high" = conf.high)
t_max_coef %>% 
  mutate(across(where(is.numeric), ~round(.x, 3))) %>% 
  flextable()
```

```{r, fig.height=12, fig.width=12}
t_max_prediction_plot <- fortify(t_max_full_m)
ggplot(t_max_prediction_plot, aes(x = .fitted, y = Tmax)) +
  geom_point(size = 2) +
  ylim(c(25,130)) +
  xlim(c(25,130)) +
  geom_abline(slope = 1, intercept = 0, color = "skyblue3",
              linetype = "dashed", linewidth = 2) +
  geom_text(x = 75, y = 130, label = "Adjusted R-squared = 0.497", size = 8) +
  labs(title = "Predicted vs observed Tmax values",
       x = "Fitted values of Tmax, min",
       y = "Observed values of Tmax, min")
```

```{r, fig.height=16, fig.width=20}
t_max_coef %>% 
ggplot(aes(y = `Variable`)) +
  geom_point(aes(x = `β estimate`), size = 3) +
  geom_errorbar(aes(xmin = `95%CI low`, xmax = `95%CI high`), width = 0.5) +
  geom_vline(xintercept = 0, color = "darkred") +
  labs(title = "Coefficients and 95%CI in Tmax model")
```

### 1.6.3. Regression models for Cmax

log(Cmax) \~ dose_g_per_kg + route + food + site

```{r Cmax-lm2-v1, fig.height=16, fig.width=16}
order_0_df_factor <- order_0_df %>% 
  mutate(
    route = factor(route),
    food = factor(food),
    site = factor(site)
  )

Cmax_lm2_v1 <- lm(log(Cmax) ~ dose_g_per_kg + route + food + site, data = order_0_df_factor)

summary(Cmax_lm2_v1)
check_model(Cmax_lm2_v1)

broom::tidy(Cmax_lm2_v1, conf.int = TRUE)
# analysis of model linearity using Residuals vs Fitted plot
autoplot(Cmax_lm2_v1, 1, ncol = 1)
# Linearity diagnosis with ggResidpanel library
resid_panel(Cmax_lm2_v1, plots = "resid", smoother = TRUE)
# Diagnosis of homoscedasticity of residuals using the Scale-Location plot
autoplot(Cmax_lm2_v1, 3, ncol = 1)
# Diagnosing the normality of residual distribution using a Normal Q-Q plot:
resid_panel(Cmax_lm2_v1, plots = c("qq", "hist"))
# Diagnosis of outliers and anomalous observations using the Residuals vs Leverage plot and Cook's distance plot
resid_panel(Cmax_lm2_v1, plots = c("lev", "cookd"))
```

log(Cmax) \~ dose_g_per_kg + route + food + site + sex + age

```{r Cmax lm2 v2, fig.height=8, fig.width=8}
order_0_df_factor <- order_0_df %>% 
  mutate(
    route = factor(route),
    food = factor(food),
    site = factor(site),
    sex = factor(sex)
  )

Cmax_lm2_v2 <- lm(log(Cmax) ~ dose_g_per_kg + route + food + sex + age, data = order_0_df_factor)

summary(Cmax_lm2_v2)
check_model(Cmax_lm2_v1)

broom::tidy(Cmax_lm2_v2, conf.int = TRUE)
# analysis of model linearity using Residuals vs Fitted plot
autoplot(Cmax_lm2_v2, 1, ncol = 1)
# Linearity diagnosis with ggResidpanel library
resid_panel(Cmax_lm2_v2, plots = "resid", smoother = TRUE)
# Diagnosis of homoscedasticity of residuals using the Scale-Location plot
autoplot(Cmax_lm2_v2, 3, ncol = 1)
# Diagnosing the normality of residual distribution using a Normal Q-Q plot:
resid_panel(Cmax_lm2_v2, plots = c("qq", "hist"))
# Diagnosis of outliers and anomalous observations using the Residuals vs Leverage plot and Cook's distance plot
resid_panel(Cmax_lm2_v2, plots = c("lev", "cookd"))
```



```{r, fig.height=10, fig.width=10}
t_max_coef <- coeftest(Cmax_lm2_v2, vcov. = vcovHC, type = "HC3") %>%
broom::tidy(conf.int = TRUE) %>% 
  mutate(term = c("Intercept",
                  "Dose (g/kg)",
                  "Peroral route",
                  "Food status: \"Fed\"",
                  "sex M",
                  "age")) %>% 
  rename("Variable" = term,
         "β estimate" = estimate,
         "Standard error" = std.error,
         "Wald t-test statistics" = statistic,
         "P-value" = p.value,
         "95%CI low" = conf.low,
         "95%CI high" = conf.high)
t_max_coef %>% 
  mutate(across(where(is.numeric), ~round(.x, 3))) %>% 
  flextable()
```


```{r, fig.height=16, fig.width=20}
ggpredict(Cmax_lm2_v2, terms = c("age", "route"), ci_level = 0) %>% plot(show_data = TRUE, log_y = FALSE, jitter = .01) + labs(subtitle = "Adjusted R-squared:  0.8459")
```


```{r, fig.height=16, fig.width=20}
plot_model(Cmax_lm2_v2, show.values = TRUE,  show.intercept = TRUE, value.offset = 0.3, digits = 3,title = "Regression coeffecients with 95% CI") +
  geom_hline(yintercept = 0, color = "gray3", size = 0.5) +
  theme_nice()
```


log(Cmax) \~ age + lbm_kg

```{r Cmax-lm2-v3, fig.height=16, fig.width=20}

Cmax_lm2_v3 <- lm(Cmax ~ age + lbm_kg, data = order_0_df_factor)

summary(Cmax_lm2_v3)
#check_model(Cmax_lm2_v1)

broom::tidy(Cmax_lm2_v3, conf.int = TRUE)
# analysis of model linearity using Residuals vs Fitted plot
autoplot(Cmax_lm2_v3, 1, ncol = 1)
# Linearity diagnosis with ggResidpanel library
resid_panel(Cmax_lm2_v3, plots = "resid", smoother = TRUE)
# Diagnosis of homoscedasticity of residuals using the Scale-Location plot
autoplot(Cmax_lm2_v3, 3, ncol = 1)
# Diagnosing the normality of residual distribution using a Normal Q-Q plot:
resid_panel(Cmax_lm2_v3, plots = c("qq", "hist"))
# Diagnosis of outliers and anomalous observations using the Residuals vs Leverage plot and Cook's distance plot
resid_panel(Cmax_lm2_v3, plots = c("lev", "cookd"))
```

### 1.6.4. Regression models for C_0

```{r С_0_0_order-lm1-v1}
order_0_df
C_0_0_lm1_v1 <- lm(log(C_0_0_order) ~ dose_g_per_kg + route + food + site + study_id, data = order_0_df)
summary(C_0_0_lm1_v1)
check_model(C_0_0_lm1_v1)

```
Adjusted R-squared:  0.8286 

1) Posterior Predictive Check

The density lines of the models and observations almost coincide. This means that the model adequately reproduces the distribution of C_0_0.

2) Linearity

The green smoothed line is close to horizontal, but there is a slight systematic trend visible at the right end (with large fitted values).
There is no significant violation of linearity. It is possible that there is insufficient interaction or nonlinear effect of one of the predictors.

3) Homogeneity of variance
There is no heteroscedasticity.

4) Influential observations

Several influential observations are visible, marked with numbers (78, 92, 62, 95, 67)

5) Normality of residuals

The Q–Q chart looks acceptable.

6) Collinearty VIF
dose_g_per_kg, route, study_id - High ≥ 10

I will try to remove unnecessary indicators, for example, in each study, 1 level of food.

```{r С_0_0_order-lm1-var1.1}
C_0_0_lm1_v2 <- lm(log(C_0_0_order) ~ study_id, data = order_0_df)
summary(C_0_0_lm1_v2)
check_model(C_0_0_lm1_v2)

```
Adjusted R-squared:  0.8277
Removed collinearity.

```{r С_0_0_order-lm2-v1-v2-part1}
order_0_df
C_0_0_lm2_v1 <- lm(log(C_0_0_order) ~ study_id + sex+age , data = order_0_df)
summary(C_0_0_lm2_v1) #sex and age depend on study_id, there is no point in looking at all the data


df_sa <- order_0_df %>% filter(!is.na(sex), !is.na(age), !is.na(C_0_0_order)) 
# Wilkinson and Thierauf data, 18 observations

df_sa <- df_sa %>%  mutate(bmi = wt_kg / (bh_cm/100)^2, 
         bsa_m2 = sqrt(bh_cm * wt_kg / 3600)) # bsa fits better in the model

C_0_0_lm2_v2 <- lm(log(C_0_0_order) ~ log(dose_g_per_kg) + bsa_m2 + sex, data = df_sa) 
summary(C_0_0_lm2_v2)
check_model(C_0_0_lm2_v2)
```

It looks better visually, but there are a lot of predictors. I calculated BMI and BSA and tried to include them in the model in various ways.

```{r С_0_0_order-lm3-v1}
C_0_0_lm3_v1 <- lm(log(C_0_0_order) ~ bsa_m2 , data = order_0_df)

summary(C_0_0_lm3_v1)
check_model(C_0_0_lm3_v1)
```

### 1.6.5. Regression models for Vd

```{r basic_lms}
fit_Vd_1 <- lm(V_d_0_order ~ route + food + site, data = order_0_df_factor)

summary(fit_Vd_1)

cat("\n===============================================================\n")

fit_Vd_3 <- lm(V_d_0_order ~ route, data = order_0_df_factor)

summary(fit_Vd_3)

cat("\n===============================================================\n")

fit_Vd_4 <- lm(V_d_0_order ~ route + site, data = order_0_df_factor)

summary(fit_Vd_4)

cat("\n===============================================================\n")

fit_Vd_5 <- lm(V_d_0_order ~ route + food, data = order_0_df_factor)

summary(fit_Vd_5)

cat("\n===============================================================\n")

anova(fit_Vd_1, fit_Vd_3)

anova(fit_Vd_1, fit_Vd_4)

anova(fit_Vd_1, fit_Vd_5)
```

```{r}
fit_Vd_3_sub <- lm(V_d_0_order ~ 1,  data = order_0_df_factor)

anova(fit_Vd_3, fit_Vd_3_sub)
```

```{r model_check, fig.width = 16, fig.height = 8}
check_model(fit_Vd_3)
```

```{r}
fit_Vd_3_add <- lm(V_d_0_order ~ route + wt_kg * sex, data = order_0_df_factor)
fit_Vd_3_add2 <- lm(V_d_0_order ~ route + wt_kg + sex, data = order_0_df_factor)

anova(fit_Vd_3_add, fit_Vd_3_add2)

summary(fit_Vd_3_add2)

fit_Vd <- lm(V_d_0_order ~ wt_kg + sex, data = order_0_df_factor)

anova(fit_Vd, fit_Vd_3_add2)

summary(fit_Vd)

```

```{r, fig.width = 16, fig.height = 8}
check_model(fit_Vd)
```

```{r}
resid_panel(fit_Vd, plots = c("lev", "cookd"))
```

```{r}
ggpredict(fit_Vd, terms = c("wt_kg", "sex"), ci_level = 0.95) %>% plot(show_data = TRUE, log_y = FALSE, jitter = .01) +
  labs(title = "Predicted values of Vd",
    subtitle = "Adjusted R-squared:  0.2259",
    color = "Sex") +
  xlab("Weight (kg)") +
  ylab("Vd (L / kg)")
```

```{r}
plot_model(fit_Vd, digits = 3, show.values = TRUE,  show.intercept = TRUE, value.offset = 0.3,  axis.labels = c("Weight (kg)", "Sex", "Intercept"), title = "Regression coeffecients with 95% CI") +
  theme_nice()
```

### 1.6.6. Regression models for Beta

```{r}
fit_Beta_1 <- lm(Beta ~ dose_g_per_kg * route, data = order_0_df_factor)

summary(fit_Beta_1)
```

```{r}
fit_Beta_1_sub <- lm(Beta ~ 1,  data = order_0_df_factor)

anova(fit_Beta_1, fit_Beta_1_sub)
```

```{r, fig.width = 16, fig.height = 8}
check_model(fit_Beta_1)
```

```{r, fig.width = 16, fig.height = 8}
resid_panel(fit_Beta_1, plots = c("lev", "cookd"))
```

```{r}
ggpredict(fit_Beta_1, terms = c("dose_g_per_kg", "route"), ci_level = 0) %>% plot(show_data = TRUE, log_y = FALSE, jitter = .01) + labs(subtitle = "Adjusted R-squared:  0.5671") + xlab("Dose (g per kg)")
```

```{r}
plot_model(fit_Beta_1, show.values = TRUE,  show.intercept = TRUE, value.offset = 0.3,  axis.labels = c("Dose (g per kg)", "Route (PO)", "Dose (g per kg) x Route (PO)", "Intercept"), title = "Regression coeffecients with 95% CI") +
  theme_nice()
```

## 1.7. Interpretation of models. Effect of factors on PK-parameters

### 1.7.1. The effect of body mass and body surface area on PK-parameters

**Body mass**

```{r}
ggpredict(fit_Vd, terms = c("wt_kg", "sex"), ci_level = 0.95) %>% plot(show_data = TRUE, log_y = FALSE, jitter = .01) +
  labs(title = "Predicted values of Vd",
    subtitle = "Adjusted R-squared:  0.2259",
    color = "Sex") +
  xlab("Weight (kg)") +
  ylab("Vd (L / kg)")
```

```{r}
fit_Vd %>% broom::tidy(conf.int = TRUE) %>% filter(term == "wt_kg") %>% 
  mutate(across(c(where(is.numeric), -p.value), ~round(.x, 3))) %>% 
  mutate(p.value = gtsummary::style_pvalue(p.value),
         term = "Body mass, kg") %>% 
  select(-statistic) %>% 
  slice_tail(n = 1) %>% 
  rename(
    `Indicator` = term,
    `Coefficient (β)` = estimate,
    `Standard error` = std.error,
    `P-value` = p.value,
    `Lower bound of 95% CI` = conf.low,
    `Upper bound of 95% CI` = conf.high
  ) %>% 
  flextable() %>%
  add_header_row(., values = "Estimation of the body mass coefficient in the Vd model", colwidth = 6) %>% 
  flex_default_fun()
```

**Body surface area**

```{r}
ggpredict(lm(V_d_0_order ~ bsa_m2 + sex, data = order_0_df),
          terms = c("bsa_m2", "sex"), ci_level = 0.95) %>% plot(show_data = TRUE, log_y = FALSE, jitter = .01) +
  labs(title = "Predicted values of Vd",
    subtitle = "Adjusted R-squared:  0.2433",
    color = "Sex") +
  xlab("Body Surface Area (m²)") +
  ylab("Vd (L / kg)")
```

```{r}
lm(V_d_0_order ~ bsa_m2 + sex, data = order_0_df) %>% 
  broom::tidy(conf.int = TRUE) %>% filter(term == "bsa_m2") %>% 
  mutate(across(c(where(is.numeric), -p.value), ~round(.x, 3))) %>% 
  mutate(p.value = gtsummary::style_pvalue(p.value),
         term = "Body surface area, m²") %>% 
  select(-statistic) %>% 
  slice_tail(n = 1) %>% 
  rename(
    `Indicator` = term,
    `Coefficient (β)` = estimate,
    `Standard error` = std.error,
    `P-value` = p.value,
    `Lower bound of 95% CI` = conf.low,
    `Upper bound of 95% CI` = conf.high
  ) %>% 
  flextable() %>%
  add_header_row(., values = "Estimation of the body surface area coefficient in the Vd model", colwidth = 6) %>% 
  flex_default_fun()
```

**Lean body mass**

```{r}
ggpredict(Cmax_lm2_v3,
          terms = c("lbm_kg", "age"), ci_level = 0.95) %>% plot(show_data = TRUE, log_y = FALSE, jitter = .01) +
  labs(title = "Predicted values of Cmax",
    subtitle = "Adjusted R-squared:  0.3328",
    color = "Age") +
  xlab("Lean Body Mass (kg)") +
  ylab("Cmax (g/L)")
```

```{r}
Cmax_lm2_v3 %>% 
  broom::tidy(conf.int = TRUE) %>%
  # filter(term == "lbm_kg") %>% 
  mutate(across(c(where(is.numeric), -p.value), ~round(.x, 4))) %>% 
  mutate(p.value = gtsummary::style_pvalue(p.value),
         term = c("Intersept", "Age, years", "Lean body mass, kg")) %>% 
  select(-statistic) %>% 
  rename(
    `Indicator` = term,
    `Coefficient (β)` = estimate,
    `Standard error` = std.error,
    `P-value` = p.value,
    `Lower bound of 95% CI` = conf.low,
    `Upper bound of 95% CI` = conf.high
  ) %>% 
  flextable() %>%
  add_header_row(., values = "Estimation of the lean body mass coefficient in the Cmax model", colwidth = 6) %>% 
  flex_default_fun()
```

With an increase in body weight by 1 kg, Vd decreases by 0.004 (95% CI [0.001; 0.007]) L/kg (all other things being equal), and with an increase in body surface area by 1 $m^2$, Vd decreases by 0.311 (95% CI [0.032; 0.589]) L/kg (all other things being equal).

When lean body mass increases by 1 kg, Cmax decreases by 0.0086 (95% CI [0.0002; 0.017]) g/L, all other things being equal.

### 1.7.2. The effect of sex on PK-parameters

Sex does not significantly affect AUC (Male 95%CI[-0.3108; 0.1754]) or Cmax (Male 95%CI[-0.1505; 0.326]).

In male subjects, the volume of distribution is 0.106 L/kg higher than in females, all other things being equal (95% CI [-0.0155; 0.1964]).

```{r}
fit_Vd %>% 
  broom::tidy(conf.int = TRUE) %>% filter(term == "sexM") %>% 
  mutate(across(c(where(is.numeric), -p.value), ~round(.x, 4))) %>% 
  mutate(p.value = gtsummary::style_pvalue(p.value),
         term = "Sex") %>% 
  select(-statistic) %>% 
  slice_tail(n = 1) %>% 
  rename(
    `Indicator` = term,
    `Coefficient (β)` = estimate,
    `Standard error` = std.error,
    `P-value` = p.value,
    `Lower bound of 95% CI` = conf.low,
    `Upper bound of 95% CI` = conf.high
  ) %>% 
  flextable() %>%
  add_header_row(., values = " Estimation of the sex coefficient in the Vd model", colwidth = 6) %>% 
  flex_default_fun()
```

### 1.7.3. The effect of food on PK-parameters

It's necessary to replace the site to food from the previous model of AUC (fit_AUC_mod_int), because you can't add them together.

```{r}
fit_AUC_mod_with_food <- lm(log(AUC) ~ dose_g_per_kg*route + food, order_0_df)

summary(fit_AUC_mod_with_food, confint = TRUE, digits = 3)
```

```{r, fig.width=6, fig.height=4, dpi=300}
plot_auc <- plot_model(fit_AUC_mod_with_food, show.values = TRUE,  show.intercept = TRUE, value.offset = 0.3, axis.labels = c("Dose (g per kg) x Route (PO)", "Food status (fed)", "Route (PO)", "Dose (g per kg)","Intercept")) +
  geom_hline(yintercept = 0, color = "gray3", size = 0.5) +
  labs(title = "Regression coefficients for AUC with 95% CI", 
       subtitle = "log(AUC) ~ dose_g_per_kg*route + food") +
  theme_nice()+
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5))

plot_auc

d <- plot_auc$data

d$ci_label <- sprintf(
  "[%.2f; %.2f]",
  d$conf.low,
  d$conf.high
)
```

Compared with the fasted state, “fed” condition was associated with log(AUC) decreased by 0.36, AUC decreased by 30.2% (all other things being equal).Percent derived as (exp(β)−1)×100.

```{r, fig.height=4, fig.width=6, dpi=300}
plot_cmax <- plot_model(Cmax_lm2_v2, show.values = TRUE,  show.intercept = TRUE, value.offset = 0.3, digits = 3) +
  geom_hline(yintercept = 0, color = "gray3", size = 0.5) +
  labs(title = "Regression coefficients for Cmax with 95% CI", 
       subtitle = "log(Cmax) ~ dose_g_per_kg + route + food + sex + age") +
  theme_nice()+
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5))

plot_cmax
```

Compared with the fasted state, “fed” condition was associated with log(Cmax) decreased by 0.286, Cmax decreased by 24.9% (all other things being equal).

```{r, fig.width = 4, fig.height = 4, dpi=300}
plot_beta <- plot_model(fit_Beta_1, show.values = TRUE,  show.intercept = TRUE, value.offset = 0.5, digits = 3,   axis.labels = c("Dose (g per kg) x Route (PO)", "food (fed)","Route (PO)", "Dose (g per kg)", "Intercept")) +
    geom_hline(yintercept = 0, color = "gray3", size = 0.5) +
labs(title = "Regression coefficients \n for Beta with 95% CI", 
       subtitle = "Beta ~ dose_g_per_kg * route + food") +
  scale_y_continuous(limits = c(-0.015, 0.015)) +
  theme_nice()+
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5))

plot_beta

d <- plot_beta$data

d$ci_label <- sprintf(
  "[%.4f; %.4f]",
  d$conf.low,
  d$conf.high)
```

Compared with the fasted state, “fed” condition was associated with Beta increased by 0.0004 g/L/min (all other things being equal). 

# 2. Allometric scaling using rats' PK parameters

## 2.1. Data unification

```{r rat_Kozawa_prep}
Kozawa_2007 <- Kozawa_2007 %>% 
  mutate(time_min = round(time_h * 60, 0)) %>% 
  mutate(route = route %>% str_remove_all(" bolus")) %>%
  select(-time_h) %>% 
  mutate(number = dose_g_per_kg/0.5)
```


```{r rat_Eriksson_prep}
Eriksson_1977 <- Eriksson_1977 %>% 
  mutate(time_min = round(time_h * 60, 0)) %>% 
  select(-time_h) %>% 
  mutate(number = case_when(
      dose_g_per_kg == 0.75 ~ 1,
      dose_g_per_kg == 1.50 ~ 2,
      dose_g_per_kg == 3.00 ~ 3,
    ))
```

## 2.2. Concentration-time plots

```{r, fig.height=12, fig.width=15, dpi=300}
ggplot(Kozawa_2007, aes(x = time_min, y = conc_g_per_L)) +
  geom_line(aes(group = number)) +
  scale_y_continuous(breaks = seq(0, 3.6, 0.2)) +
  scale_x_continuous(breaks = seq(0, 500, 25)) +
  labs(title = "Concentration-time plot (rats, IV route)",
       subtitle = "Averaged data",
       caption = "Source: Kozawa-2007",
       x = "Time, min",
       y = "Concentration, g/L")
```


```{r, fig.height=12, fig.width=15, dpi=300}
ggplot(Eriksson_1977, aes(x = time_min, y = conc_g_per_L)) +
  geom_line(aes(group = number)) +
  scale_y_continuous(breaks = seq(0, 1.8, 0.1)) +
  scale_x_continuous(breaks = seq(0, 300, 30)) +
  labs(title = "Concentration-time plot (rats, PO route)",
       subtitle = "Averaged data",
       caption = "Source: Eriksson-1977",
       x = "Time, min",
       y = "Concentration, g/L")
```

## 2.3. PK-parameteres calculation

```{r rat_Kozawa_pk}
Kozawa_2007_pk <- Kozawa_2007 %>% 
  arrange(time_min) %>% 
  nest(.by = number, .key = "average_data") %>% 
  mutate(pharma = purrr::map(average_data,
                            ~pharma_fun(.$conc_g_per_L,
                                        .$time_min,
                                        .$dose_g_per_kg))) %>% 
  unnest_wider(col = pharma) %>% 
  select(-average_data)

Kozawa_2007_pk <- Kozawa_2007 %>% 
  select(-c(time_min, conc_g_per_L)) %>% 
  group_by(number) %>% 
  slice(1) %>% 
  ungroup() %>% 
  right_join(Kozawa_2007_pk) %>% 
  select(-number)
```


```{r rat_Eriksson_pk}
Eriksson_1977_pk <- Eriksson_1977 %>% 
  arrange(time_min) %>% 
  nest(.by = number, .key = "average_data") %>% 
  mutate(pharma = purrr::map(average_data,
                            ~pharma_fun(.$conc_g_per_L,
                                        .$time_min,
                                        .$dose_g_per_kg))) %>% 
  unnest_wider(col = pharma) %>% 
  select(-average_data)

# for two concentrations there are not enough points to calculate beta

Eriksson_1977_pk <- Eriksson_1977 %>% 
  select(-c(time_min, conc_g_per_L)) %>% 
  group_by(number) %>% 
  slice(1) %>% 
  ungroup() %>% 
  right_join(Eriksson_1977_pk) %>% 
  select(-number)
```

```{r rat-join}
rat_data_pk <- rbind(Kozawa_2007_pk, Eriksson_1977_pk)
```

## 2.4. Preaparing data and function

```{r rat_data} 
# Dataset prepararion
rat_data_pk <-  rat_data_pk %>% 
  mutate(V_max_0_order = Beta * 60 * V_d_0_order,
         BW_g = BW_g / 1000) %>%
  rename(wt_kg = BW_g) %>% 
  relocate(V_max_0_order, .after = V_d_0_order)
```

```{r allometric_func}
allom_scale <- function(V_rat, # L, abs
                        rat_weight, # kg
                        human_weight = 1, # kg
                        n = 1)
{
  V_human = V_rat * (human_weight / rat_weight) ^ n
  return(V_human)
}
```

```{r allometric}
# Human dataset preparation
allometric <- order_0_df %>% 
  filter(!is.na(wt_kg)) %>% 
  select(study_id, subject_id, dose_g_per_kg, route, 
         Beta, V_d_0_order, wt_kg) %>% 
  mutate(V_max = Beta * V_d_0_order * 60) %>% 
  mutate(Vd_obs = V_d_0_order * wt_kg,
         Vmax_obs = Beta * V_d_0_order * wt_kg * 60,
         Beta_obs = Beta * 60) %>% 
  relocate(c(Vd_obs, Vmax_obs, Beta_obs), .after = V_max)

# Only Wilkinson IV (Human dose 0,5-0,62 mg/kg)
# allometric_IV_Wilk <- allometric %>% 
#   filter(route == "IV",
#          study_id == "Wilkinson-1976-IV") %>% 
#   crossing(rat_data_pk %>% filter(dose_g_per_kg == 0.5) %>% 
#              mutate(Vd_abs_rat = (V_d_0_order * wt_kg)) %>% 
#              rename_with(~ paste0("rat_", .x))) %>% 
#   mutate(Vd_classic = allom_scale(rat_Vd_abs_rat, rat_wt_kg, wt_kg),
#          Vd_matsumoto = allom_scale(rat_Vd_abs_rat, rat_wt_kg, wt_kg, 0.932),
#          Beta_classic = allom_scale(rat_Beta, rat_wt_kg, wt_kg, 0.75),
#          Beta_matsumoto = allom_scale(rat_Beta, rat_wt_kg, wt_kg, 0.708),
#          Vmax_classic = allom_scale(rat_V_max_0_order, rat_wt_kg, wt_kg, 0.75),
#          Vmax_matsumoto = allom_scale(rat_V_max_0_order, rat_wt_kg, wt_kg, 0.708))
```

## 2.5. Allometric scaling

## 2.5.1. IV allometric scaling (main df)

```{r allometric_IV}
# Allometric data for IV route
allometric_IV <- allometric %>% 
  filter(route == "IV") %>% 
  crossing(rat_data_pk %>% filter(dose_g_per_kg == 0.5) %>% 
             mutate(Vd_abs_rat = (V_d_0_order * wt_kg)) %>% 
             rename_with(~ paste0("rat_", .x))) %>% 
  mutate(Vd_classic = allom_scale(rat_Vd_abs_rat, rat_wt_kg, wt_kg),
         Vd_matsumoto = allom_scale(rat_Vd_abs_rat, rat_wt_kg, wt_kg, 0.932),
         Beta_classic = allom_scale(rat_Beta * 60, rat_wt_kg, wt_kg, 0.75 - 1),
         Beta_matsumoto = allom_scale(rat_Beta * 60, rat_wt_kg, wt_kg, 0.708 - 0.932),
         Vmax_classic = allom_scale(rat_V_max_0_order, rat_wt_kg, wt_kg, 0.75),
         Vmax_matsumoto = allom_scale(rat_V_max_0_order, rat_wt_kg, wt_kg, 0.708)) %>% 
  mutate(Vd_metr_cl = (Vd_classic - Vd_obs) / Vd_obs * 100,
         Vd_metr_mats = (Vd_matsumoto - Vd_obs) / Vd_obs * 100,
         Beta_metr_cl = (Beta_classic - Beta_obs) / Beta_obs * 100,
         Beta_metr_mats = (Beta_matsumoto - Beta_obs) / Beta_obs * 100,
         Vmax_metr_cl = (Vmax_classic - Vmax_obs) / Vmax_obs * 100,
         Vmax_metr_mats = (Vmax_matsumoto - Vmax_obs) / Vmax_obs * 100
  ) %>% 
  mutate(across(contains("metr"), ~round(.x, 3)))
```

```{r main_dose_IV}
# Observed and mean dose
main_df_IV <- order_0_df %>% 
  filter(route == "IV" & !is.na(wt_kg)) %>% 
  mutate(Vd_obs = V_d_0_order * wt_kg,
         Vmax_obs = Beta * V_d_0_order * wt_kg * 60,
         Beta_obs = Beta * 60)

cat("Human mean dose, IV route, g/kg:", mean(main_df_IV$dose_g_per_kg))
```

```{r Vd_prop_IV, fig.width=5, fig.height=3, dpi=300}
allometric_IV %>% 
  select(contains("Vd_metr")) %>% 
  rename(`Allometric classic` = Vd_metr_cl,
         `Matsumoto` = Vd_metr_mats) %>% 
  pivot_longer(everything(), names_to = "Type", values_to = "Percentage") %>% 
  filter(!is.na(Percentage)) %>% 
ggplot(aes(x = Type, y = Percentage)) +
  geom_boxplot(width = 0.1, staplewidth = 0.5) +
  labs(
    x = "Used allometric scale exponent",
    y = "Error, %", 
    title = "Error of Vd allometric scaling using different allometric\nscale exponents, IV route of administration",
    subtitle = "Rat dose = 0.5 g/kg"
  ) +
  theme_bw()
```

```{r Vd_visual_IV}
# Vd
Vd_vis_IV <- bind_rows(
  main_df_IV %>% transmute(value = Vd_obs, model = "Observed"),
  allometric_IV %>% transmute(value = Vd_classic, model = "Allometric classic"),
  allometric_IV %>% transmute(value = Vd_matsumoto, model = "Matsumoto")
)

ggplot(Vd_vis_IV, aes(x = model, y = value)) +
  geom_boxplot(width = 0.1, staplewidth = 0.5) +
  theme_bw()+
  labs(
    x = "Used allometric scale exponent",
    y = "Vd, L", 
    title = "Vd, allometric scaling using different allometric\nscale exponents, IV route of administration",
    subtitle = "Mean human dose = 0.57 g/kg, rat dose = 0.5 g/kg"
  )
```

The usage of classical exponent (=1) for Vd allometric scaling shows mean error (%) from +20 to +50%. Matsumoto's exponent (=0,932) shows mean error (%) from 0% to -15%.

```{r Beta_prop_IV, fig.width=6, fig.height=4, dpi=300}
allometric_IV %>% 
  select(contains("Beta_metr")) %>% 
  rename(`Allometric classic` = Beta_metr_cl,
         `Matsumoto` = Beta_metr_mats) %>% 
  pivot_longer(everything(), names_to = "Type", values_to = "Percentage") %>% 
  filter(!is.na(Percentage)) %>% 
ggplot(aes(x = Type, y = Percentage)) +
  geom_boxplot(width = 0.1, staplewidth = 0.5) +
  labs(
    x = "Used allometric scale exponent",
    y = "Error, %", 
    title = "Error of Beta allometric scaling using different allometric\nscale exponents, IV route of administration",
    subtitle = "Rat dose = 0.5 g/kg"
  ) +
  theme_bw()
```

```{r Beta_visual_IV}
# Beta
Beta_vis_IV <- bind_rows(
  main_df_IV %>% transmute(value = Beta_obs, model = "Observed"),
  allometric_IV %>% transmute(value = Beta_classic, model = "Allometric classic"),
  allometric_IV %>% transmute(value = Beta_matsumoto, model = "Matsumoto")
)

ggplot(Beta_vis_IV, aes(x = model, y = value)) +
  geom_boxplot(width = 0.1, staplewidth = 0.5) +
  theme_bw()+
  labs(
    x = "Used allometric scale exponent",
    y = "Beta, g/L/hour", 
    title = "Beta, allometric scaling using different allometric\nscale exponents, IV route of administration",
    subtitle = "Mean human dose = 0.57 g/kg, rat dose = 0.5 g/kg"
  )

```

The usage of classical exponent (= 0,75 - 1) for Beta allometric scaling shows mean error (%) from -35% to -50%. Matsumoto's exponent (= 0,708 - 0,932) shows mean error (%) from -25% to -43%.

```{r Vmax_prop_IV, fig.width=6, fig.height=4, dpi=300}
allometric_IV %>% 
  select(contains("Vmax_metr")) %>% 
  rename(`Allometric classic` = Vmax_metr_cl,
         `Matsumoto` = Vmax_metr_mats) %>% 
  pivot_longer(everything(), names_to = "Type", values_to = "Percentage") %>% 
  filter(!is.na(Percentage)) %>% 
ggplot(aes(x = Type, y = Percentage)) +
  geom_boxplot(width = 0.1, staplewidth = 0.5) +
  labs(
    x = "Used allometric scale exponent",
    y = "Error, %", 
    title = "Error of Vmax allometric scaling using different allometric\nscale exponents, IV route of administration",
    subtitle = "Rat dose = 0.5 g/kg"
  ) +
  theme_bw()
```

```{r Vmax_visual_IV}
# Vmax
Vmax_vis_IV <- bind_rows(
  main_df_IV %>% transmute(value = Vmax_obs, model = "Observed"),
  allometric_IV %>% transmute(value = Vmax_classic, model = "Allometric classic"),
  allometric_IV %>% transmute(value = Vmax_matsumoto, model = "Matsumoto")
)

ggplot(Vmax_vis_IV, aes(x = model, y = value)) +
  geom_boxplot(width = 0.1, staplewidth = 0.5) +
  theme_bw()+
  labs(
    x = "Used allometric scale exponent",
    y = "Vmax, g/hour", 
    title = "Vmax, allometric scaling using different allometric\nscale exponents, IV route of administration",
    subtitle = "Mean human dose = 0.57 g/kg, rat dose = 0.5 g/kg"
  )
```

The usage of classical exponent (= 0,75) for Vmax allometric scaling shows mean error (%) from +87 to +150%. Matsumoto's exponent (= 0,708) shows mean error (%) from +50% to +100%.

### 2.5.2. PO allometric scaling

#### 2.5.2.1. PO allometric scaling (750 mg per kg)

```{r allometric_PO_750_mg}
# Allometric scaling
allometric_PO <- allometric %>% 
  filter(route == "PO") %>% 
  crossing(rat_data_pk %>% filter(dose_g_per_kg == 0.75) %>% 
             mutate(Vd_abs_rat = (V_d_0_order * wt_kg)) %>% 
             rename_with(~ paste0("rat_", .x))) %>% 
  mutate(Vd_classic = allom_scale(rat_Vd_abs_rat, rat_wt_kg, wt_kg),
         Vd_matsumoto = allom_scale(rat_Vd_abs_rat, rat_wt_kg, wt_kg, 0.932)) %>% 
  mutate(Vd_metr_cl = (Vd_classic - Vd_obs) / Vd_obs * 100,
         Vd_metr_mats = (Vd_matsumoto - Vd_obs) / Vd_obs * 100,
  ) %>% 
  mutate(across(contains("metr"), ~round(.x, 3)))
  
```

```{r main_dose_PO}
# Human dataset preparation
main_df_PO <- order_0_df %>% 
  filter(route == "PO" & !is.na(wt_kg)) %>% 
  mutate(Vd_obs = V_d_0_order * wt_kg,
         Vmax_obs = Beta * V_d_0_order * wt_kg * 60,
         Beta_obs = Beta * 60)

cat("Human mean dose, PO route, g/kg):", mean(main_df_PO$dose_g_per_kg))
```

```{r Vd_prop_PO_750, fig.width=6, fig.height=4, dpi=300}
allometric_PO %>% 
  select(contains("Vd_metr")) %>% 
  rename(`Allometric classic` = Vd_metr_cl,
         `Matsumoto` = Vd_metr_mats) %>% 
  pivot_longer(everything(), names_to = "Type", values_to = "Percentage") %>% 
  filter(!is.na(Percentage)) %>% 
ggplot(aes(x = Type, y = Percentage)) +
  geom_boxplot(width = 0.1, staplewidth = 0.5) +
  labs(
    x = "Used allometric scale exponent",
    y = "Error, %", 
    title = "Error of Vd allometric scaling using different allometric\nscale exponents, PO route of administration",
    subtitle = "Rat dose = 0.75 g/kg"
  ) +
  theme_bw()
```

Mean error (%) for Vd allometric scaling is more than +100% up to +300% using classic allometric exponent.
We cannot use another PK parameters due to no captured elimination phase in the rat data for this dose.

```{r Vd_visual_PO_750_mg}
# Vd
Vd_vis_PO <- bind_rows(
  main_df_PO %>% transmute(value = Vd_obs, model = "Observed"),
  allometric_PO %>% transmute(value = Vd_classic, model = "Allometric classic"),
  allometric_PO %>% transmute(value = Vd_matsumoto, model = "Matsumoto")
)

ggplot(Vd_vis_PO, aes(x = model, y = value)) +
  geom_boxplot(width = 0.1, staplewidth = 0.5) +
  theme_bw()+
  labs(
    x = "Used allometric scale exponent",
    y = "Vd, L", 
    title = "Vd, allometric scaling using different allometric\nscale exponents, PO route of administration",
    subtitle = "Mean human dose = 0.58 g/kg, rat dose = 0.75 g/kg"
  )
```

#### 2.5.2.2. PO allometric scaling (3000 mg per kg)

```{r allometric_PO_3_g}
# Scaling
allometric_PO_3 <- allometric %>% 
  filter(route == "PO") %>% 
  crossing(rat_data_pk %>% filter(dose_g_per_kg == 3) %>% 
             mutate(Vd_abs_rat = (V_d_0_order * wt_kg)) %>% 
             rename_with(~ paste0("rat_", .x))) %>%
  mutate(Vd_classic = allom_scale(rat_Vd_abs_rat, rat_wt_kg, wt_kg),
         Vd_matsumoto = allom_scale(rat_Vd_abs_rat, rat_wt_kg, wt_kg, 0.932),
         Beta_classic = allom_scale(rat_Beta * 60, rat_wt_kg, wt_kg, 0.75 - 1),
         Beta_matsumoto = allom_scale(rat_Beta * 60, rat_wt_kg, wt_kg, 0.708 - 0.932),
         Vmax_classic = allom_scale(rat_V_max_0_order, rat_wt_kg, wt_kg, 0.75),
         Vmax_matsumoto = allom_scale(rat_V_max_0_order, rat_wt_kg, wt_kg, 0.708)) %>% 
  mutate(Vd_metr_cl = (Vd_classic - Vd_obs) / Vd_obs * 100,
         Vd_metr_mats = (Vd_matsumoto - Vd_obs) / Vd_obs * 100,
         Beta_metr_cl = (Beta_classic - Beta_obs) / Beta_obs * 100,
         Beta_metr_mats = (Beta_matsumoto - Beta_obs) / Beta_obs * 100,
         Vmax_metr_cl = (Vmax_classic - Vmax_obs) / Vmax_obs * 100,
         Vmax_metr_mats = (Vmax_matsumoto - Vmax_obs) / Vmax_obs * 100
  ) %>% 
  mutate(across(contains("metr"), ~round(.x, 3)))
  
```

```{r Vd_prop_PO_3000, fig.width=6, fig.height=4, dpi=300}
allometric_PO_3 %>% 
  select(contains("Vd_metr")) %>% 
  rename(`Allometric classic` = Vd_metr_cl,
         `Matsumoto` = Vd_metr_mats) %>% 
  pivot_longer(everything(), names_to = "Type", values_to = "Percentage") %>% 
  filter(!is.na(Percentage)) %>% 
ggplot(aes(x = Type, y = Percentage)) +
  geom_boxplot(width = 0.1, staplewidth = 0.5) +
  labs(
    x = "Used allometric scale exponent",
    y = "Error, %", 
    title = "Error of Vd allometric scaling using different allometric\nscale exponents, PO route of administration",
    subtitle = "Rat dose = 3.00 g/kg"
  ) +
  theme_bw()
```

```{r Vd_visual_PO_3_g}
# Vd
Vd_vis_PO_3 <- bind_rows(
  main_df_PO %>% transmute(value = Vd_obs, model = "Observed"),
  allometric_PO_3 %>% transmute(value = Vd_classic, model = "Allometric classic"),
  allometric_PO_3 %>% transmute(value = Vd_matsumoto, model = "Matsumoto")
)

ggplot(Vd_vis_PO_3, aes(x = model, y = value)) +
  geom_boxplot(width = 0.1, staplewidth = 0.5) +
  theme_bw()+
  labs(
    x = "Used allometric scale exponent",
    y = "Vd, L", 
    title = "Vd, allometric scaling using different allometric\nscale exponents, PO route of administration",
    subtitle = "Mean human dose = 0.58 g/kg, rat dose = 3.00 g/kg"
  )
```

```{r Beta_prop_PO_3000, fig.width=6, fig.height=4, dpi=300}
allometric_PO_3 %>% 
  select(contains("Beta_metr")) %>% 
  rename(`Allometric classic` = Beta_metr_cl,
         `Matsumoto` = Beta_metr_mats) %>% 
  pivot_longer(everything(), names_to = "Type", values_to = "Percentage") %>% 
  filter(!is.na(Percentage)) %>% 
ggplot(aes(x = Type, y = Percentage)) +
  geom_boxplot(width = 0.1, staplewidth = 0.5) +
  labs(
    x = "Used allometric scale exponent",
    y = "Error, %", 
    title = "Error of Beta allometric scaling using different allometric\nscale exponents, PO route of administration",
    subtitle = "Rat dose = 3.00 g/kg"
  ) +
  theme_bw()
```

```{r Beta_visual_PO_3_g}
# Beta
Beta_vis_PO_3 <- bind_rows(
  main_df_PO %>% transmute(value = Beta_obs, model = "Observed"),
  allometric_PO_3 %>% transmute(value = Beta_classic, model = "Allometric classic"),
  allometric_PO_3 %>% transmute(value = Beta_matsumoto, model = "Matsumoto")
)

ggplot(Beta_vis_PO_3, aes(x = model, y = value)) +
  geom_boxplot(width = 0.1, staplewidth = 0.5) +
  theme_bw()+
  labs(
    x = "Used allometric scale exponent",
    y = "Beta, g/L/hour", 
    title = "Beta, allometric scaling using different allometric\nscale exponents, PO route of administration",
    subtitle = "Mean human dose = 0.58 g/kg, rat dose = 3.00 g/kg"
  )
```

```{r Vmax_prop_PO_3000, fig.width=6, fig.height=4, dpi=300}
allometric_PO_3 %>% 
  select(contains("Vmax_metr")) %>% 
  rename(`Allometric classic` = Vmax_metr_cl,
         `Matsumoto` = Vmax_metr_mats) %>% 
  pivot_longer(everything(), names_to = "Type", values_to = "Percentage") %>% 
  filter(!is.na(Percentage)) %>% 
ggplot(aes(x = Type, y = Percentage)) +
  geom_boxplot(width = 0.1, staplewidth = 0.5) +
  labs(
    x = "Used allometric scale exponent",
    y = "Error, %", 
    title = "Error of Vmax allometric scaling using different allometric\nscale exponents, PO route of administration",
    subtitle = "Rat dose = 3.00 g/kg"
  ) +
  theme_bw()
```

```{r Vmax_visual_PO_3_g}
# Vmax
Vmax_vis_PO_3 <- bind_rows(
  main_df_PO %>% transmute(value = Vmax_obs, model = "Observed"),
  allometric_PO_3 %>% transmute(value = Vmax_classic, model = "Allometric classic"),
  allometric_PO_3 %>% transmute(value = Vmax_matsumoto, model = "Matsumoto")
)

ggplot(Vmax_vis_PO_3, aes(x = model, y = value)) +
  geom_boxplot(width = 0.1, staplewidth = 0.5) +
  theme_bw()+
  labs(
    x = "Used allometric scale exponent",
    y = "Vmax, g/hour", 
    title = "Vmax, allometric scaling using different allometric\nscale exponents, PO route of administration",
    subtitle = "Mean human dose = 0.58 g/kg, rat dose = 3.00 g/kg"
  )
```

Predictive mean performance (error, %) for PO route with rat dose 3.0 g/kg: 

- Vd: Allometric classic +187.82, Matsumoto +98.65;


- Beta: Allometric classic -72.20, Matsumoto -67.97;


- Vmax: Allometric classic +144.25, Matsumoto +94.28.